On the Bit Complexity of Finding Points in Connected Components of a Smooth Real Hypersurface

### 2020-05-15 main.tex
Updates from Overleaf

This being said, our two key results are the following.
\begin{theorem}\label{theo:gen}
  For $i=1,\dots,n$, there exists a non-zero polynomial $\D_i$ in
  $\C[\A]$ of degree at most $6n^2(2d)^{2n}$ such that if $\mA \in
  \C^{n\times n}$ does not cancel $\D_i$, then $\mA$ is invertible and
  $f^\mA$ satisfies $\textbf{H}_i$.
\end{theorem}

@@

\section{Property ${\bf H}_i(2)$: Noether position}\label{sec:Hi2}
Let $f\in \ZZ[X_1,\hdots,X_n]$, with total degree $d$ and $V(f)\subset \C^n$ smooth, and $i \in \{1,\hdots,n\}$ be fixed throughout this section. We
prove that there exists a non-zero polynomial $\D_{i}$ in $\C[\A]$ of degree at most $6n^2(2d)^{2n}$ such that if $\mA \in \C^{n \times n}$ does not cancel $\D_{i}$,
then $\mA$ is invertible and $f^{\mA}$ satisfies both conditions in
$\textbf{H}_i$.

@@

We saw in Section~\ref{ssec:Hi1} that $\D_{i,1}$ has degree at most
$2nd^{2n}$, and all $D_j$'s have degree at most $(2d)^n$. This gives
the upper bound
$$\deg(\D_i) \le 2nd^{2n} + 2n^2(2d)^{2n} + n(2d)^n \le 6n^2 (2d)^{2n}.$$
This completes the proof of Theorem~\ref{theo:gen}.

@@

\nl Construct $S := \{1,2,\hdots,\lceil 3\epsilon^{-1}6n^3(2d)^{2n} \rceil\}$ and $T :=
\{1,2,\hdots,\lceil 3\epsilon^{-1}nd^{2n}\rceil\},$ and randomly choose $\mA \in
S^{n^2}$, and $\bm \sigma \in T^{n-1}$\;

@@

\paragraph*{Bit operation cost:} The following lists the costs for each step of Algorithm 1:
\begin{enumerate}
    \item We defined $S := \{1,2,\hdots,\lceil 3\epsilon^{-1}6n^3(2d)^{2n} \rceil \}$
      and therefore the height of any $a_{i,j} \in S$ is at most
    \[
    \log 3/{\epsilon} + \log(6n^3(2d)^{2n}) \in O^{\sim}(\log 1/{\epsilon} + n\log d).
    \]
    \item 
    After computing the partial derivatives, the height grows by at
    most another factor of $\log d$. Thus, all polynomials in the
    system considered at Step 3 have height $ O^{\sim}(b + d\log 1 /
    \epsilon + dn)$. All integer coefficients appearing in $\Gamma_i$ 
    satisfy the same bound.    
    %% Furthermore, preparing the
    %% straight line program for the gradient of $f$ does not introduce
    %% large integers. The integers in $\grad (f)$ are of the same
    %% magnitude as in $\Gamma$.
    \item As a result, after applying \cite[Algorithm 2]{SH} $k$ times
      for each index $i$, with $k = O(\log n + \log 1 / \epsilon)$,
      the total boolean cost of the algorithm is
      \[
      O^{\sim}(d^{3n+1}(\log1/\epsilon)(b + \log1/\epsilon))
      \]
      where the polynomials in the output have degree at most $d^n,$ and height at most
      \[
      O^{\sim}(d^{n+1}(b + \log 1 / \epsilon)).
  \]
\end{enumerate}

@@

\paragraph*{Probability of success.} 
Let $\Delta_i \in \C[\A]$ be the polynomials from
Theorem~\ref{theo:gen}. Denote by $\Delta := \prod_{i=1}^n \D_i,$ and
note that
\begin{align}
    \deg \Delta \leq \sum_{i=1}^n \deg \Delta_i \leq 6n^3(2d)^{2n}.
\end{align}
If $\mA \in \C^{n \times n}$ does not cancel $\Delta,$ then $\mA$ is
invertible and $f^\mA$ satisfies $\textbf{H}_i$ for all $i \in
\{1,\hdots,n\}.$ Now, assuming that $\mA$ is such a matrix, let
$\Xi_i\in \C[\A]$ be the polynomials from Theorem 2.2 applied to
$f^{\mA}.$ Denote by $\Xi := \prod_{i=1}^n \Xi_i,$ and note that
\begin{align}
    \deg \Xi \leq \sum_{i=1}^n \deg \Xi_i \leq nd^{2n}.
\end{align}
If $\bm \sigma \in \C^{i-1}$ does not cancel $\Xi$, then $f^{\mA}$ and
$\bm \sigma$ satisfy $\textbf{H}_i^{'}$ for all $i \in
\{1,\hdots,n\};$ as we argued above, the algorithm is guaranteed to
succeed, as long as our call to Algorithm 2 in~\cite{SH} succeeds. That latter reference establishes that by repeating the calculation $k$ times, and keeping the output of highest degree among those $k$ results, we succeed with probability at least $1-(1/2)^k$. When Algorithm 2 does not succeed, it either returns a proper subset of the solutions, or FAIL. Note that Algorithm 2 is shown to succeed in a single run with  probability at least $1-11/32,$ and we bound the probability of success with $1-1/2$ for simplicity. 
Now, by construction of
\[
S := \{1,2,\hdots,\lceil 3\epsilon^{-1}6n^3(2d)^{2n}\rceil \}
\] 
and 
\[
T := \{1,2,\hdots,\lceil 3\epsilon^{-1}nd^{2n} \rceil \},
\] 
where $\mA \in S^{n^2}$ and  $\bm \sigma \in T^{n-1}$ are randomly chosen, we have 
%
\[
\pr[\Delta(\mA)=0] \leq  \frac{\deg\Delta}{|S|} = \epsilon/3
\]
and
\[
\pr[\Xi(\bm \sigma)=0] \leq  \frac{\deg\Xi}{|T|} = \epsilon/3.
\]
%
Let $\mathscr{E}$ be the event that the parametrizations
$[\mathscr{Q}_1,\hdots,\mathscr{Q}_n]$ returned in step 6 of
Algorithm~1 are correct. Then, the probability of success is equal to
\begin{align*}
 \pr[\Delta(\mA)\ne 0] \times \pr[\Xi(\bm \sigma)\ne 0 ~|~\Delta(\mA) \ne 0] \times \pr[\mathscr{E}~|~
   \Delta(\mA)\Xi(\bm\sigma) \ne 0].
\end{align*}
%
Set $k = \lg(3n/\epsilon)$ so that 
\[
(1-2^{-k})^n = (1 - \epsilon/(3n))^n \geq 1 - \epsilon/3,
\]
by Bernoulli's inequality. Therefore, 
\begin{align*}
\pr[\textrm{success}] &\geq (1- \epsilon/3)(1-
\epsilon/3)\pr[\mathscr{E}~|~   \Delta(\mA)\Xi(\bm\sigma) \ne 0]\\ 
& \geq (1- \epsilon/3)(1-
\epsilon/3)(1-2^{-k})^n \\ & \geq (1- \epsilon/3)(1- \epsilon/3)(1-
\epsilon/3)\\ & \geq 1 - \epsilon.
\end{align*}
This finishes the proof of our main theorem.

>>>>

This being said, our two key results are the following.
\begin{theorem}\label{theo:gen}
  For $i=1,\dots,n$, there exists a non-zero polynomial $\D_i$ in
  $\C[\A]$ of degree at most $5n^2(2d)^{2n}$ such that if $\mA \in
  \C^{n\times n}$ does not cancel $\D_i$, then $\mA$ is invertible and
  $f^\mA$ satisfies $\textbf{H}_i$.
\end{theorem}

@@

\section{Property ${\bf H}_i(2)$: Noether position}\label{sec:Hi2}
Let $f\in \ZZ[X_1,\hdots,X_n]$, with total degree $d$ and $V(f)\subset \C^n$ smooth, and $i \in \{1,\hdots,n\}$ be fixed throughout this section. We
prove that there exists a non-zero polynomial $\D_{i}$ in $\C[\A]$ of degree at most $5n^2(2d)^{2n}$ such that if $\mA \in \C^{n \times n}$ does not cancel $\D_{i}$,
then $\mA$ is invertible and $f^{\mA}$ satisfies both conditions in
$\textbf{H}_i$.

@@

We saw in Section~\ref{ssec:Hi1} that $\D_{i,1}$ has degree at most
$2nd^{2n}$, and all $D_j$'s have degree at most $(2d)^n$. This gives
the upper bound
$$\deg(\D_i) \le 2nd^{2n} + 2n^2(2d)^{2n} + n(2d)^n \le 5n^2 (2d)^{2n}.$$
This completes the proof of Theorem~\ref{theo:gen}.

@@

\nl Construct $S := \{1,2,\hdots,\lceil 3\epsilon^{-1}5n^3(2d)^{2n} \rceil\}$ and $T :=
\{1,2,\hdots,\lceil 3\epsilon^{-1}nd^{2n}\rceil\},$ and randomly choose $\mA \in
S^{n^2}$, and $\bm \sigma \in T^{n-1}$\;

@@

\paragraph*{Bit operation cost:} The following lists the costs for each step of Algorithm 1:
\begin{enumerate}
    \item We defined $S := \{1,2,\hdots,\lceil 3\epsilon^{-1}5n^3(2d)^{2n} \rceil \}$
      and therefore the height of any $a_{i,j} \in S$ is at most
    \[
    \log 3/{\epsilon} + \log(5n^3(2d)^{2n}) \in O^{\sim}(\log 1/{\epsilon} + n\log d).
    \]
    \item 
    After computing the partial derivatives, the height grows by at
    most another factor of $\log d$. Thus, all polynomials in the
    system considered at Step 3 have height $ O^{\sim}(b + d\log 1 /
    \epsilon + dn)$. All integer coefficients appearing in $\Gamma_i$ 
    satisfy the same bound.    
    %% Furthermore, preparing the
    %% straight line program for the gradient of $f$ does not introduce
    %% large integers. The integers in $\grad (f)$ are of the same
    %% magnitude as in $\Gamma$.
    \item As a result, after applying \cite[Algorithm 2]{SH} $k$ times
      for each index $i$, with $k = O(\log n + \log 1 / \epsilon)$,
      the total boolean cost of the algorithm is
      \[
      O^{\sim}(d^{3n+1}(\log1/\epsilon)(b + \log1/\epsilon))
      \]
      where the polynomials in the output have degree at most $d^n,$ and height at most
      \[
      O^{\sim}(d^{n+1}(b + \log 1 / \epsilon)).
  \]
\end{enumerate}

@@

\paragraph*{Probability of success.} 
Let $\Delta_i \in \C[\A]$ be the polynomials from
Theorem~\ref{theo:gen}. Denote by $\Delta := \prod_{i=1}^n \D_i,$ and
note that
\begin{align}
    \deg \Delta \leq \sum_{i=1}^n \deg \Delta_i \leq 5n^3(2d)^{2n}.
\end{align}
If $\mA \in \C^{n \times n}$ does not cancel $\Delta,$ then $\mA$ is
invertible and $f^\mA$ satisfies $\textbf{H}_i$ for all $i \in
\{1,\hdots,n\}.$ Now, assuming that $\mA$ is such a matrix, let
$\Xi_i\in \C[\A]$ be the polynomials from Theorem 2.2 applied to
$f^{\mA}.$ Denote by $\Xi := \prod_{i=1}^n \Xi_i,$ and note that
\begin{align}
    \deg \Xi \leq \sum_{i=1}^n \deg \Xi_i \leq nd^{2n}.
\end{align}
If $\bm \sigma \in \C^{i-1}$ does not cancel $\Xi$, then $f^{\mA}$ and
$\bm \sigma$ satisfy $\textbf{H}_i^{'}$ for all $i \in
\{1,\hdots,n\};$ as we argued above, the algorithm is guaranteed to
succeed, as long as our call to Algorithm 2 in~\cite{SH} succeeds. That latter reference establishes that by repeating the calculation $k$ times, and keeping the output of highest degree among those $k$ results, we succeed with probability at least $1-(1/2)^k$. When Algorithm 2 does not succeed, it either returns a proper subset of the solutions, or FAIL. Note that Algorithm 2 is shown to succeed in a single run with  probability at least $1-11/32,$ and we bound the probability of success with $1-1/2$ for simplicity. 
Now, by construction of
\[
S := \{1,2,\hdots,\lceil 3\epsilon^{-1}5n^3(2d)^{2n}\rceil \}
\] 
and 
\[
T := \{1,2,\hdots,\lceil 3\epsilon^{-1}nd^{2n} \rceil \},
\] 
where $\mA \in S^{n^2}$ and  $\bm \sigma \in T^{n-1}$ are randomly chosen, we have 
%
\[
\pr[\Delta(\mA)=0] \leq  \frac{\deg\Delta}{|S|} = \epsilon/3
\]
and
\[
\pr[\Xi(\bm \sigma)=0] \leq  \frac{\deg\Xi}{|T|} = \epsilon/3.
\]
%
Let $\mathscr{E}$ be the event that the parametrizations
$[\mathscr{Q}_1,\hdots,\mathscr{Q}_n]$ returned in step 6 of
Algorithm~1 are correct. Then, the probability of success is equal to
\begin{align*}
 \pr[\Delta(\mA)\ne 0] \times \pr[\Xi(\bm \sigma)\ne 0 ~|~\Delta(\mA) \ne 0] \times \pr[\mathscr{E}~|~
   \Delta(\mA)\Xi(\bm\sigma) \ne 0].
\end{align*}
%
Set $k = \lg(3n/\epsilon)$ so that 
\[
(1-2^{-k})^n = (1 - \epsilon/(3n))^n \geq 1 - \epsilon/3,
\]
by Bernoulli's inequality. Therefore, 
\begin{align*}
\pr[\textrm{success}] &\geq (1- \epsilon/3)(1-
\epsilon/3)\pr[\mathscr{E}~|~   \Delta(\mA)\Xi(\bm\sigma) \ne 0]\\ 
& \geq (1- \epsilon/3)(1-
\epsilon/3)(1-2^{-k})^n \\ & \geq (1- \epsilon/3)(1- \epsilon/3)(1-
\epsilon/3)\\ & \geq 1 - \epsilon.
\end{align*}
This finishes the proof of our main theorem.
